
In this section we present an evaluation of the parameterised verification
procedures described in Section~\ref{sec:verification} on the guarding game
presented in Example~\ref{ex:agent-template}.


%%%%
The guarding game is an instance of a social dilemma game, a class of
MASs characterised by tension between individual
and collective rationality \cite{VanlangeJPV13}. The  game simulates
the fundamental forces of a \emph{collective risk dilemma (CRD)}, a type of
social dilemma where a guaranteed ``tragedy of the commons''
\cite{Hardin68} is avoided by personal sacrifice by a population of
agents, or brought on by free riding if all the agents %favour a strategy that
neglect the collective interests % of that personal sacrifice
\cite{SantosP11}.
%
Namely, guarding can be considered equivalent to cooperation (acting in
collective interest), and resting to defection (acting in selfish
interest).


We train a neural observation function using deep Q-learning, a type of
reinforcement learning (RL) algorithm. During the training, the game was played
by 4 agents, and the parameters were set as $M_h = 4$, $G_r = -2$, $R_r = 1$
and $U_r = -3$. The rewards were assigned to reflect the tension between
individual and collective interests. All agents shared the same neural network,
and thus
% creating a system in which the agents
were learning to play against exact copies of themselves.
%
\nb{E: commented out experience replay buffer and target network}% We used an experience replay buffer and a separate target network which was
% updated only periodically to stabilise overestimation of the Q-values
% \cite{Mnih+15,HaaseltGS16}.
% More details about the training can be found in the Appendix.
%
The produced neural network has two hidden layers of four ReLU activated
neurons, takes as input a single neuron, representing the normalised health
points of the agent, and outputs the estimated Q-values of the two actions
`rest' and `guard'.
%%%%


\begin{table}
\centering
% \begin{tabular}{c@{\qquad}rrrrr}
%   \toprule
%   k &\multicolumn{1}{c}{$n=2$} & \multicolumn{1}{c}{$n=3$} & \multicolumn{1}{c}{$n=4$} & \multicolumn{1}{c}{$n=5$} & \multicolumn{1}{c}{$n=6$}\\\midrule
  
%   2 &              0.09s &     0.13s &   0.53s &    1.15s &              5.09s\\
%   3 & \graycell    1.46s &     0.30s &   1.19s &    3.41s &             17.58s\\
%   4 & \graycell    5.49s &     0.52s &   2.31s &   17.74s & \multicolumn{1}{c}{--}\\
%   5 & \graycell   61.47s &   133.28s &   4.28s &   95.83s & \multicolumn{1}{c}{--}\\
%   \bottomrule
% \end{tabular}
\begin{tabular}{c@{\quad}rrrr}
  \toprule
  &
    \begin{tabular}{@{}c@{}}
      $k=2$\\[-0.5mm] ($n=2$)
    \end{tabular}
  &
    \begin{tabular}{@{}c@{}}
      $k=3$\\[-0.5mm] ($n=3$)
    \end{tabular}
  &
    \begin{tabular}{@{}c@{}}
      $k=4$\\[-0.5mm] ($n=3$)
    \end{tabular}
  &
    \begin{tabular}{@{}c@{}}
      $k=5$\\[-0.5mm] ($n=3$)
    \end{tabular}
  \\\midrule
  
  $i=2$ &  0.09s & \graycell  1.46s & \graycell  5.49s & \graycell  61.47s \\
  $i=3$ &  0.13s &            0.30s &            0.52s &           133.28s \\[1mm]
  $i=4$ &  0.53s &            1.19s &            2.31s &             4.28s \\
  $i=5$ &  1.15s &            3.41s &           17.74s &            95.83s \\
  $i=6$ &  5.09s &           17.58s & \multicolumn{1}{c}{--} & \multicolumn{1}{c}{--} \\
  \bottomrule
\end{tabular}
\caption{ Verification times for $\msys{i} \models \varphi^k_E[m]$ for $m=2$
  and various $k$ and $i$. \textcolor{red}{For each $k$, we indicate the value of $n$ from
  Corollary~\ref{cor:existential}.} Grey cells indicate when the property was
  not satisfied.  Dashes indicate a 1 hour timeout.  }
  \label{tab:results-existential}
\end{table}

Given the learned neural network, we implemented a template agent and a
zero-one agent for the guarding game. We then used the \venmas toolkit
developed for verification of NIS~\cite{Akintunde+20b} to verify whether it is
possible for a colony of agents to survive after a number of time steps.
%
Specifically, recall from Example~\ref{ex:pnis} that proposition~$\mathsf{a}$
labels all states with positive health (``alive'') and proposition $\mathsf{d}$
labels all other states (``dead'').
%
Let $m$ be the number of agent variables. We considered two
specifications (for $v_i \in\atvar$):
\begin{inparaenum}[\it (i)]
\item the existential property
  $\varphi^k_E = EX^k \bigwedge_{i=1}^m(\mathsf{a},v_i)$ 
  expressing that there is an evolution where at least $m$ agents are alive after
  $k$ time steps,
\item the universal property $\varphi^k_A = AX^k \bigwedge_{i=1}^m(\mathsf{a},v_i)$
  expressing that in every possible evolution at least $m$ agents are alive
  after $k$ time steps.
\end{inparaenum}
%
The experiments %pertaining to the verification of these properties
were performed on a standard PC running Ubuntu 22.04 with 16GB RAM and
processor Intel(R) Core i5-4460. We relied on Gurobi
v10.0~\cite{Gurobi+16a} to solve the mixed integer linear programs generated by
\venmas.

\nb{E: the whole paragraph has been updated}%
Let $\pnis$ be the parameterised NIS for the guarding example. For the
existential property, we consider the case of $m=2$ and are interested in
checking whether at least two agents can stay alive for $k$ time steps
($\pnis\models\forall_{v_1,v_2}\varphi^k_E$), or whether there is a minimal
number (an emergence threshold) of agents that is required to guarantee
that. We vary the temporal depth~$k$ from 2 till 5.  To verify if
$\pnis\models\forall_{v_1,v_2}\varphi^k_E$, according to
Corollary~\ref{cor:existential} we need to check whether
$\msys{i}\models\forall_{v_1,v_2}\varphi^k_E$ for every $i\in\{m,\dots,n\}$,
where $n$ is the minimal $i$ such that $\msys{i}$ $k$-bounded simulates
$\masys{m}$. In our case $n=2$ for $k=2$ and $n=3$ for $k\geq 3$.  In
Table~\ref{tab:results-existential} we present the outcomes of the verification
queries $\msys{i}\models\forall_{v_1,v_2}\varphi^k_E$ for $i=2,3$, as well as
for $i=4,5,6$. The latter are not used to reason about parameterised
verification; instead, they are included to demonstrate the increasing
computational cost of verifying concrete systems for increasing number of
agents, empirically justifying the need for parameterised verification.
%
For temporal depth $k=2$, we have that
$\pnis\models\forall_{v_1,v_2}\varphi^k_E$. This is no longer the case for at
least 3 time steps as the system composed of two agents violates the property:
$\msys{2}\not\models\forall_{v_1,v_2}\varphi^k_E$. On the other hand, we obtain
that $n=3$ is the emergence threshold for $\forall_{v_1,v_2}\varphi^k_E$ when
$k\geq 3$, that is, there need to be at least 3 agents for the colony to be
viable.

For the universal property, we verified $\varphi^k_A[m]$ against
\textcolor{red}{the abstract model} $\masys{m}$ for temporal depth
$k = 1,\dots,6$ and the number of agents $m=2,3$. The verification times and
results can be found in Table~\ref{tab:results-universal}. We observe that due
to the presence of the zero-one agent, the verification times are longer than
the ones observed in Table~\ref{tab:results-existential}. We additionally
notice that the property is not satisfied even after 1 time step, which is
expected given that there paths where no agent is guarding even when there are
volunteers.

\nb{E: new}Our experimental results confirm the intractability of verification for
concrete models as the number of agents grows, thereby motivating the need for
the parameterised verification methods that we put forward.





% The arbitrary selection of a subset of guards that volunteer is a necessary
% artefact of PNIS. Therefore, the evaluation of a universal property like
% $\varphi^k_A[m]$ will consider paths where nobody is guarding even when there
% are volunteers. Nevertheless, we are able to reason about existential
% properties in a meaningful way, as the only event which would incur the tragedy
% of the commons is the lack of any volunteer guarding in the first place.
% % The arbitrary selection of a subset of guards that volunteer is a
% % necessary artefact of the interpreted systems, but does not affect the
% % stability of the system negatively and allows us to reason about its
% % existential properties, as the only event which would incur the tragedy of the
% % commons is the lack of any volunteer guards in the first place.


\begin{table}
  \centering
  \begin{tabular}{@{}crrrrrr@{}}
    \toprule
    % $k$ & \multicolumn{1}{c}{$m = 2$} & \multicolumn{1}{c}{$m = 3$}\\
    % \midrule
    % 1 & \graycell        0.66s      & \graycell        1.75s\\
    % 2 & \graycell        3.60s      & \graycell       16.03s\\
    % 3 & \multicolumn{1}{c}{--}      & \graycell        0.00s\\
    % 4 & \graycell       16.90s      & \multicolumn{1}{c}{--}\\
    % 5 & \graycell       20.59s      & \multicolumn{1}{c}{--}\\
    % 6 & \multicolumn{1}{c}{--}      & \multicolumn{1}{c}{--}\\

    m & \multicolumn{1}{c}{$k = 1$} & \multicolumn{1}{c}{$k = 2$} & \multicolumn{1}{c}{$k = 3$} & \multicolumn{1}{c}{$k = 4$} & \multicolumn{1}{c}{$k = 5$} & \multicolumn{1}{c}{$k = 6$}\\
    \midrule
2 & \graycell    0.66s & \graycell    3.60s & \multicolumn{1}{c}{--} & \graycell   16.90s & \graycell   20.59s & \multicolumn{1}{c}{--}\\
3 & \graycell    1.75s & \graycell   16.03s & \multicolumn{1}{c}{--} & \graycell 42.74s & \graycell 2196.70s & \multicolumn{1}{c}{--} \\
    \bottomrule
  \end{tabular}
  \caption{ Verification times for $\masys{m} \models \varphi^k_A[m]$ for various
    $m$ and $k$.  Grey cells indicate when the property was not satisfied.
    Dashes indicate a 1 hour timeout.  }
  \label{tab:results-universal}
\end{table}



%%% Local Variables:
%%% mode: latex
%%% fill-column: 79
%%% TeX-master: "../main"
%%% End:
